---
title: "Misinformation Analysis"
author: "Daniel Allington"
date: "2025-03-06"
output:
  word_document: default
  html_document: default
---

```{r setup}
library(dplyr)
library(readr)
library(ggplot2)

options(scipen = 999)

misinfo_d <- read_csv(
  '../data/life-under-lockdown-wave2-extract.csv',
  show_col_types = FALSE
)

knowledge_levels <- c(
  'Nothing at all',
  'Not very much',
  'A fair amount',
  'A great deal'
)

misinfo_d <- misinfo_d %>%
  mutate(
    know_facebook = know_facebook %>%
      ordered(
        levels = knowledge_levels
      )
  )

```

# Descriptive statistics
```{r visualise-gender}
misinfo_d %>%
  ggplot(
    aes(x = gender)
  ) +
  geom_bar()
```

```{r visualise-age}
misinfo_d %>%
  ggplot(
    aes(x = age)
  ) +
  geom_histogram(binwidth = 5)
```

```{r visualise-facebook-knowledge}
misinfo_d %>%
  ggplot(
    aes(x = know_facebook)
  ) +
  geom_bar()
```

```{r visualise-beliefs}
misinfo_d %>%
  ggplot(
    aes(x = beliefs)
  ) +
  geom_histogram(binwidth = 1)
```

# Hypotheses
H1 We expect a relationship between gender and number of misinformation beliefs.

H2 We expect a correlation between age and number of misinformation beliefs.

H3 We expect a correlation between amount of knowledge gained from Facebook and number of misinformation beliefs.

# Findings
## Gender versus number of misinformation beliefs
```{r gender-beliefs-table}
misinfo_d %>%
  group_by(
    gender
  ) %>%
  summarise(
    n = n(),
    M = mean(beliefs),
    Mdn = median(beliefs),
    SD = sd(beliefs)
  )
```

```{r}
misinfo_d %>%
  filter(gender != 'In another way') %>%
  t.test(
    beliefs ~ gender,
    data = .,
    var.equal = FALSE
  )
```

In the sample, men had more misinformation beliefs (M = 0.73, SD = 1.13) than women (M = 0.61, SD = 1.05), but this was not statistically significant, t(494.31) = -1.201, p = .230.

## Age versus number of misinformation beliefs
```{r age-beliefs-visualisation}
misinfo_d %>%
  ggplot(aes(x = age, y = beliefs)) +
  geom_jitter()
```

```{r age-beliefs-correlation}
misinfo_d %>%
  cor.test(
    ~ age + beliefs,
    data = .,
    method = 'pearson'
  )
```

In the sample, there is a statistically significant moderate negative correlation between age and number of misinformation beliefs, r(498) = -.24, 95% CI [-.32, -.16], p < .001.

## Knowledge from Facebook versus number of misinformation beliefs
```{r facebook-beliefs-visualisation}
misinfo_d %>%
  ggplot(aes(x = know_facebook, y = beliefs)) +
  geom_boxplot()
```

```{r facebook-beliefs-correlation}
misinfo_d %>%
  mutate(
    know_facebook = as.numeric(know_facebook)
  ) %>%
  cor.test(
    ~ know_facebook + beliefs,
    data = .,
    method = 'spearman',
    exact = FALSE
  )
```

In the sample, there is a moderate positive correlation between amount of knowledge gained from Facebook and number of misinformation beliefs, rs(498) = .23, p < .001.



## Linear model
```{r create-dummy-variables}
model_d <- misinfo_d %>%
  mutate(
    Female = as.numeric(gender == 'Female'),
    High_Facebook = as.numeric(know_facebook %in% c('A fair amount', 'A great deal'))
  )
```

```{r age-model}
model_d %>%
  lm(
    beliefs ~ age,
    data = .
  ) %>%
  summary
```

```{r facebook-model}
model_d %>%
  lm(
    beliefs ~ High_Facebook,
    data = .
  ) %>%
  summary
```

```{r gender-model}
model_d %>%
  lm(
    beliefs ~ Female,
    data = .
  ) %>%
  summary
```

```{r full-model}
full_model <- model_d %>%
  lm(
    beliefs ~ age + High_Facebook + Female,
    data = .
  )

full_model %>%
  summary
```


